---
title: "CCISS Loss/Gain Stats"
author: "Kiri Daust"
format: html
editor: visual
---

# Community Stats

Connect to full duckdb database:

```{r}
devtools::load_all()
library(data.table)
library(terra)
library(ggplot2)
library(duckdb)

con <- dbCon_cciss("./bc_2km.duckdb", threads = 8)
```

## Portfolio Persistence

Here, we define persistence at a given location as the proportion of species that have maintained persistence, from the reference period to the "future". We consider both the **establishment** future, and the **maturity** future, using the same weighted average calculations as in CCISS. We calculate persistence by cell, and can then map persistence, as well as average over the province (or by district).

```{r}
estab_wt <- c(0.25,0.25,0.5,0,0,0)
maturity_wt <- c(0,0,0.1,0.3,0.3,0.3)

dat <- spp_persistance(con, fut_wt = maturity_wt, curr_wt = c(1,0,0,0,0,0))
dat <- dat[TotalSpp > 1,]
dat[,Persistence := StableSpp/TotalSpp]
```

Let's plot it. We're seeing low persistence in the southern valleys, and in the Northeast.

```{r}
dem2 <- rast("dem_BC2kmGrid.tif")
X <- copy(dem2)
values(X) <- NA
X[dat$SiteRef] <- dat$Persistence
terra::plot(X)
```

Now let's average by district:

```{r}
districts <- dbGetQuery(con, "select * from dist_points") |> as.data.table()
dat[districts, district := i.district, on = c(SiteRef = "cellnum")]
temp <- dat[,.(Persistence = mean(Persistence)), by = district]
setorder(temp, Persistence)
temp
```

Then, we can say something like "Average maturity portfolio persistence province wide is `r mean(dat$Persistence) * 100` %, although some districts have a `r min(temp$Persistence) * 100`% average".

## Species Richness

We define the change in species richness as the number of suitable species in the future period divided by the suitable species in the reference period. For example, if a given location currently has 6 suitable species, and in the future period it has 9 suitable species, that corresponds to a change in richness of 150%. Again, we can map this statistic:

```{r}
dat[,Richness := (FutureSpp/TotalSpp)*100]

X <- copy(dem2)
values(X) <- NA
X[dat$SiteRef] <- dat$Richness

brks <- c(seq(0, 100, by = 10),
          seq(125, 400, by = 25))

# Colours: losses -> blue/purple, neutral -> white, gains -> orange/red
cols_loss <- hcl.colors(sum(brks < 100), "BluYl", rev = TRUE)
cols_gain <- hcl.colors(sum(brks > 100) - 1, "YlOrRd")

cols <- c(cols_loss, "grey", cols_gain)

terra::plot(
  X,
  breaks = brks,
  col = cols,
  range = c(0, 400),
  plg = list(title = "% of current richness"),
  fill_range = TRUE
)
```

# Per Species Stats

```{r}
spp_lg <- spp_loss_gain_temporal(con)
ref_area <- spp_lg[FuturePeriod == "1961_1990", .(Spp, RefArea = Stable)]
spp_lg[ref_area, RefArea := i.RefArea, on = "Spp"]
spp_lg[,`:=`(Loss = Loss/RefArea, Gain = Gain/RefArea, Stable = Stable/RefArea)]
spp_lg[,RefArea := NULL]
spp_lg[,Period := as.numeric(substr(FuturePeriod, 1,4)) + 9]

lg <- melt(spp_lg, id.vars = c("Spp", "FuturePeriod", "Period"), variable.name = "Stat")

ggplot(lg, aes(x = Period, y = value, col = Stat, group = Stat)) +
  geom_line() +
  facet_wrap(~Spp) +
  ylab("Proportion of Reference Area") +
  xlab("Period")

#ggsave("Loss_Gain_Figs.png", width = 10, height = 8, dpi = 400)
```
