---
title: "CCISS Loss/Gain Stats"
author: "Kiri Daust"
format: html
editor: visual
---

# Community Stats

Connect to full duckdb database:

```{r}
devtools::load_all()
library(data.table)
library(terra)
library(ggplot2)
library(duckdb)

con <- dbCon_cciss("./bc_2km.duckdb", threads = 8)
```

## Portfolio Persistence

Here, we define persistence at a given location as the proportion of species that have maintained persistence, from the reference period to the "future". We consider both the **establishment** future, and the **maturity** future, using the same weighted average calculations as in CCISS. We calculate persistence by cell, and can then map persistence, as well as average over the province (or by district).

```{r}
estab_wt <- c(0.25,0.25,0.5,0,0,0)
maturity_wt <- c(0,0,0.1,0.3,0.3,0.3)

dat <- spp_persistance(con, fut_wt = maturity_wt, curr_wt = c(1,0,0,0,0,0))
dat <- dat[TotalSpp > 0,]
dat[,Persistence := StableSpp/TotalSpp]
```

Let's plot it. We're seeing low persistence in the southern valleys, and in the Northeast.

```{r}
dem2 <- rast("dem_BC2kmGrid.tif")
X <- copy(dem2)
values(X) <- NA
X[dat$SiteRef] <- dat$Persistence
terra::plot(X)
```

Now let's average by district:

```{r}
districts <- dbGetQuery(con, "select * from dist_points") |> as.data.table()
dat[districts, district := i.district, on = c(SiteRef = "cellnum")]
temp <- dat[,.(Persistence = mean(Persistence)), by = district]
setorder(temp, Persistence)
temp
```

Then, we can say something like "Average maturity portfolio persistence province wide is `r mean(dat$Persistence) * 100` %, although some districts have a `r min(temp$Persistence) * 100`% average".

## Species Richness
